ESRC SDAI Application - Understanding domestic abuse using Big Data

### 2019-07-08 case_for_support.tex
Anna update

These two datasets have different merits, and combining them allows us to gain a deeper understanding the characteristics and dynamics of domestic abuse, and deliver policy-relevant insights. A unique benefit of the CSEW is that it provides information on a control group (those who have not suffered domestic abuse),  allowing us to investigate the causal effect of domestic abuse. In addition, the CSEW is more likely to give a better estimate of the true prevalence of domestic abuse, and contains rich demographic, socio-economic and geographic (LSOA level) information on the respondent (survivor). However, owing to the fact that it is a victimisation survey, it does not contain any information on the perpetrators, and has limited longitudinal information (respondents are asked about whether they suffered domestic abuse in the past 12 months, but the dates of those incidents are not specified). In contrast, the WMP crime dataset has information (age, gender, ethnicity) on both the victim and the offender, and records all reported incidents for the same victim-offender pair, allowing us to conduct a quantitative analysis of the dynamics of abusive relationships, and identify the predictors of escalation. Information on the exact timing of the abuse and the time of reporting will help us identify high-risk times of the year and understand the factors influencing the decision to report. Furthermore, the crime dataset provides us with information on other criminal behaviours of the offender. In all the analysis plans outlined below, we will not link the individuals in these datasets.

@@

\subsection*{1. Understanding domestic abuse: survivor characteristics and police mis-recording}

@@

\begin{table}[!htbp]
\caption{Understanding domestic abuse: survivor characteristics and police mis-recording, analysis plan}
  \begin{threeparttable}[t]
  \centering
       \begin{tabular}{ L{2.8cm}  C{1.8cm}  C{1.8cm}  C{3.2cm}  C{2.8cm}  C{2cm} }
    \toprule
     \textbf{Research question} & \textbf{Dataset}    & \textbf{Unit of analysis} & \textbf{Outcome variables} & \textbf{Explanatory variables} & \textbf{Model} \\
    \midrule
    What are the characteristics of domestic abuse victims? & CSEW & Individual (respondent-level) & Domestic abuse victimisation; Victimisation by type (physical, sexual abuse, threats) & demographic and socio-economic characteristics & Logistic regression/random forest \\
        \midrule
    How does the police response depend on victim characteristics? & CSEW (reported domestic abuse, by type of offence) & Individual (respondent-level) & Police action taken (no action, warning, arrest, charge) & demographic and socio-economic characteristics of the victim & Multinomial logistic regression with  fixed effects/ mixed effects random forest \\
     \midrule
    What predicts victims' trust in the police? & CSEW (reported domestic abuse, by type of offence) & Individual (respondent-level)& Satisfaction with police action, perception of security after reporting & police action taken, demographic and socio-economic characteristics of the victim & Multinomial logistic regression with fixed effects/ mixed effects random forest \\
%    Beznau II & 23.10.1971 & Pressurised water reactor & 365   & 2031 \\
%    Mühleberg & 01.07.1971 & Boiling water reactor & 373   & 2019\\
%    Gösgen & 02.02.1979 & Pressurised water reactor & 1010  & 2039 \\
%    Leibstadt & 24.05.1984 & Boiling water reactor & 1220  & 2044 \\
     \bottomrule
  \end{tabular}
%     \begin{tablenotes}
%     \item[1] Assumption: life time of 60 years.
%     \item[2] Official shut-down.
%   \end{tablenotes}
    \end{threeparttable}%
  \label{tab:addlabel}%
\end{table}%
\FloatBarrier
%Second, we can validate some of these findings using the WMP data, since the CSEW contains information on whether the survivor reported the domestic abuse to the police. The WMP contains information about the sex and age of victims. We can contrast the sex-age victimisation numbers from the WMP with corresponding population estimates to validate the findings of the CSEW \textit{(we were initially talking about using victims of burglary, but maybe this is better?)}. We can also complement this analysis with data on the offender's sex  and age, to estimate the prevalence of different forms of reported domestic abuse (child to parent violence, parent to child violence, intimate partner violence) and contrast it with estimates based on self-reported data from the CSEW, which distinguishes between family and intimate partner violence.

>>>>

These two datasets have different merits, and combining them allows us to gain a deeper understanding the characteristics and dynamics of domestic abuse, and deliver policy-relevant insights. A unique benefit of the CSEW is that it allows us to investigate the causal effect of domestic abuse by providing information on a control group (those who have not suffered domestic abuse). In addition, the CSEW is more likely to give a better estimate of the true prevalence of domestic abuse, and contains rich demographic, socio-economic and geographic (LSOA level) information on the respondent (survivor). However, owing to the fact that it is a victimisation survey, it does not contain any information on the perpetrators, and has limited longitudinal information (respondents are asked about whether they suffered domestic abuse in the past 12 months, but the dates of those incidents are not specified). In contrast, the WMP crime dataset has information (age, gender, ethnicity) on both the victim and the offender, and records all reported incidents for the same victim-offender pair, allowing us to conduct a quantitative analysis of the dynamics of abusive relationships, and identify the predictors of escalation. Information on the exact timing of the abuse and the time of reporting will help us identify high-risk times of the year and understand the factors influencing the decision to report. Furthermore, the crime dataset provides us with information on other criminal behaviours of the offender. In all the analysis plans outlined below, we will not link the individuals in these datasets.

@@

\subsection*{1. Exploring the characteristics of survivors, and understanding the police response to reported cases}

@@

\begin{table}[!htbp]
\caption{Exploring the characteristics of survivors, and understanding the police response to reported cases, analysis plan}
  \begin{threeparttable}[t]
  \centering
       \begin{tabular}{ L{2.8cm}  C{1.8cm}  C{1.8cm}  C{3.2cm}  C{2.8cm}  C{2cm} }
    \toprule
     \textbf{Research question} & \textbf{Dataset}    & \textbf{Unit of analysis} & \textbf{Outcome variables} & \textbf{Explanatory variables} & \textbf{Model} \\
    \midrule
    What are the characteristics of domestic abuse victims? & CSEW & Individual (respondent-level) & Domestic abuse victimisation; Victimisation by type (physical, sexual abuse, threats) & demographic and socio-economic characteristics & Logistic regression/random forest \\
        \midrule
    How does the police response depend on victim characteristics? & CSEW (reported domestic abuse, by type of offence) & Individual (respondent-level) & Police action taken (no action, warning, arrest, charge) & demographic and socio-economic characteristics of the victim & Multinomial logistic regression with  fixed effects/ mixed effects random forest \\
     \midrule
    What predicts victims' trust in the police? & CSEW (reported domestic abuse, by type of offence) & Individual (respondent-level)& Satisfaction with police action, perception of security after reporting & police action taken, demographic and socio-economic characteristics of the victim & Multinomial logistic regression with fixed effects/ mixed effects random forest \\
%    Beznau II & 23.10.1971 & Pressurised water reactor & 365   & 2031 \\
%    Mühleberg & 01.07.1971 & Boiling water reactor & 373   & 2019\\
%    Gösgen & 02.02.1979 & Pressurised water reactor & 1010  & 2039 \\
%    Leibstadt & 24.05.1984 & Boiling water reactor & 1220  & 2044 \\
     \bottomrule
  \end{tabular}
%     \begin{tablenotes}
%     \item[1] Assumption: life time of 60 years.
%     \item[2] Official shut-down.
%   \end{tablenotes}
    \end{threeparttable}%
  \label{tab:addlabel}%
\end{table}%
\FloatBarrier
%Second, we can validate some of these findings using the WMP data, since the CSEW contains information on whether the survivor reported the domestic abuse to the police. The WMP contains information about the sex and age of victims. We can contrast the sex-age victimisation numbers from the WMP with corresponding population estimates to validate the findings of the CSEW \textit{(we were initially talking about using victims of burglary, but maybe this is better?)}. We can also complement this analysis with data on the offender's sex  and age, to estimate the prevalence of different forms of reported domestic abuse (child to parent violence, parent to child violence, intimate partner violence) and contrast it with estimates based on self-reported data from the CSEW, which distinguishes between family and intimate partner violence.
